---
title: "Lab 02 - Plastic waste"
author: "Rowan Kemmerly"
date: "1.24.23"
output: github_document
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
plastic_waste <- read.csv("data/plastic-waste.csv")
```

## Exercises

### Exercise 1

First I'm looking at the distribution of plastic waste per capita in 2010. 

```{r plastic-waste-distribution}
ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap)) +
  geom_histogram(binwidth = 0.2)
```

Trinidad and Tobago seems to have a much higher level of plastic waste per capita than other countries:

```{r plastic-waste outlier}
plastic_waste %>%
  filter(plastic_waste_per_cap > 3.5)
```

It seems like this could have been an error, though??
https://tt.loopnews.com/content/swmcol-forbes-article-data-used-was-incorrect-and-dated



---------------------

*Reminder of the basic structure of data visualization with ggplot:*

*ggplot(data = [dataset],* 
     *mapping = aes(x = [x-variable],* 
                   *y = [y-variable])) +*
   *geom_xxx() +*
   *other options*

---------------------

Now I am examining plastic waste per capita separated out by continent.

```{r plastic-waste-continent}
ggplot(data = plastic_waste, 
     mapping = aes(x = plastic_waste_per_cap)) +
   geom_histogram() +
   facet_wrap(~ continent)
```

Based on these graphs, it seems like people in South America and Oceania generate less plastic waste than other continents. However, it's hard to tell based on these graphs how much plastic people produce relative to the total population of each continent — these graphs mostly tell us that a lot of people live in Africa and Asia but it's hard to distinguish each continent's distribution with this bin size.

-----------------------

Here's the same data used for the very first graph, but instead of a histogram in the form of a density plot.

```{r plastic-waste-density-1}
ggplot(data = plastic_waste,
       aes(x = plastic_waste_per_cap)) +
  geom_density()
```

Here's a density plot with density curves for each continent.

```{r plastic-waste-density-2}
ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     color = continent)) +
  geom_density()
```

Here's that same plot with each of the curves filled in with the corresponding color.

```{r plastic-waste-density-3}
ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     color = continent, 
                     fill = continent)) +
  geom_density()
```

And here's the same plot with more transparent density curve fills:

```{r plastic-waste-density-4}
ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     color = continent, 
                     fill = continent)) +
  geom_density(alpha = 0.7)
```


### Exercise 2

#### 2.1

And finally, here's a density plot that has color fills that are transparent enough to actually see what all of the curves clustered to the left side look like:

```{r plastic-waste-density-right-alpha}
ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     color = continent, 
                     fill = continent)) +
  geom_density(alpha = 0.4)
```

#### 2.2
##### Describe why we defined the color and fill of the curves by mapping aesthetics of the plot but we defined the alpha level as a characteristic of the plotting geom.

Things that we want to apply to individual variables (but not all variables) go with the "aesthetics" subfunction, while things that we want to apply to the entire graph (i.e., to all of the plotted variables) go outside of the ggplot function. The first type of specification is called "mapping" while the second type of specification is called "setting."



### Exercise 3

Remove this text, and add your answer for Exercise 3 here.

### Exercise 4

Remove this text, and add your answer for Exercise 4 here.

```{r plastic-waste-violin}
# insert code here
```

### Exercise 5

Remove this text, and add your answer for Exercise 5 here.

```{r plastic-waste-mismanaged}
# insert code here
```

### Exercise 6

Remove this text, and add your answer for Exercise 6 here.

```{r plastic-waste-mismanaged-continent}
# insert code here
```

### Exercise 7

Remove this text, and add your answer for Exercise 7 here.

```{r plastic-waste-population-total}
# insert code here
```

```{r plastic-waste-population-coastal}
# insert code here
```

### Exercise 8

Remove this text, and add your answer for Exercise 8 here.

```{r recreate-viz}
# insert code here
```


