---
title: "Lab 02 - Plastic waste"
author: "Rowan Kemmerly"
date: "1.24.23"
output: github_document
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
plastic_waste <- read.csv("data/plastic-waste.csv")
```

## Exercises

### Exercise 1

First I'm looking at the distribution of plastic waste per capita in 2010. 

```{r plastic-waste-distribution}
ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap)) +
  geom_histogram(binwidth = 0.2)
```

Trinidad and Tobago seems to have a much higher level of plastic waste per capita than other countries:

```{r plastic-waste outlier}
plastic_waste %>%
  filter(plastic_waste_per_cap > 3.5)
```

It seems like this could have been an error, though??
https://tt.loopnews.com/content/swmcol-forbes-article-data-used-was-incorrect-and-dated


In order to be able to examine the data without this outlier in the future (if I want to) I will create a subset of the data without it:

```{r plastic-waste subset}
plastic_waste_subset<- subset(plastic_waste, plastic_waste_per_cap < 3.5, 
                              select=code:total_pop)
```



---------------------

*Reminder of the basic structure of data visualization with ggplot:*

*ggplot(data = [dataset],* 
     *mapping = aes(x = [x-variable],* 
                   *y = [y-variable])) +*
   *geom_xxx() +*
   *other options*

---------------------

Now I am examining plastic waste per capita separated out by continent.

```{r plastic-waste-continent}
ggplot(data = plastic_waste, 
     mapping = aes(x = plastic_waste_per_cap)) +
   geom_histogram() +
   facet_wrap(~ continent)
```

Based on these graphs, it seems like people in South America and Oceania generate less plastic waste than other continents. However, it's hard to tell based on these graphs how much plastic people produce relative to the total population of each continent — these graphs mostly tell us that a lot of people live in Africa and Asia but it's hard to distinguish each continent's distribution with this bin size.

-----------------------

Here's the same data used for the very first graph, but instead of a histogram in the form of a density plot.

```{r plastic-waste-density-1}
ggplot(data = plastic_waste,
       aes(x = plastic_waste_per_cap)) +
  geom_density()
```

Here's a density plot with density curves for each continent.

```{r plastic-waste-density-2}
ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     color = continent)) +
  geom_density()
```

Here's that same plot with each of the curves filled in with the corresponding color.

```{r plastic-waste-density-3}
ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     color = continent, 
                     fill = continent)) +
  geom_density()
```

And here's the same plot with more transparent density curve fills:

```{r plastic-waste-density-4}
ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     color = continent, 
                     fill = continent)) +
  geom_density(alpha = 0.7)
```


### Exercise 2

#### 2.1

And finally, here's a density plot that has color fills that are transparent enough to actually see what all of the curves clustered to the left side look like:

```{r plastic-waste-density-right-alpha}
ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     color = continent, 
                     fill = continent)) +
  geom_density(alpha = 0.4)
```

#### 2.2
##### Describe why we defined the color and fill of the curves by mapping aesthetics of the plot but we defined the alpha level as a characteristic of the plotting geom.

Things that we want to apply to individual variables (but not all variables) go with the "aesthetics" subfunction, while things that we want to apply to the entire graph (i.e., to all of the plotted variables) go outside of the ggplot function. The first type of specification is called "mapping" while the second type of specification is called "setting."


### Exercise 3

Now, let's try looking at the data in the form of box plots.

```{r plastic-waste-box-plots}
ggplot(data = plastic_waste, 
       mapping = aes(x = continent, 
                     y = plastic_waste_per_cap)) +
  geom_boxplot()
```

#### 3.1

As a comparison, here is the data graphed via violin plots. 

```{r plastic-waste-violin}
ggplot(data = plastic_waste, 
       mapping = aes(x = continent, 
                     y = plastic_waste_per_cap)) +
  geom_violin()
```

Unlike the box plots, the violin plots show the shape of each distribution (i.e., it's easier to get a sense of how many cases correspond with each quantity of plastic waste). However, unlike the violin plots, the box plots clearly indicate where the interquartile range and median are.

### Exercise 4

#### 4.1

Here is yet another way to look at the data: a scatterplot!

This scatterplot depicts the relationship between plastic waste per capita and mismanaged plastic waste per capita.

```{r plastic-waste-scatterplot}
ggplot(data = plastic_waste,
       aes(x = plastic_waste_per_cap,
       y = mismanaged_plastic_waste_per_cap)) +
  geom_point()
```

It seems that there is a slight positive correlation between the two variables.

#### 4.2

Adding in color differentiation by continent shows us that it seems like there is a stronger correlation between plastic waste per capita and mismanaged plastic waste per capita in North America, Asia, and Europe than there is in Africa and Oceania. 

```{r plastic-waste-scatterplot-color}
ggplot(data = plastic_waste,
       aes(x = plastic_waste_per_cap,
       y = mismanaged_plastic_waste_per_cap,
       color = continent)) +
  geom_point()
```

#### 4.3

Here I am examining the relationship between plastic waste per capita and total population, as well as the relationship between plastic waste per capita and coastal population. I decided to look at the data both with and without the Trinidad and Tobago outlier because I was having a hard time getting a good sense of the data with the outlier included (as this clumps the data all in the bottom left corner).

```{r plastic-waste-population-total}
ggplot(data = plastic_waste,
       aes(x = plastic_waste_per_cap,
       y = total_pop)) +
  geom_point()
```

```{r plastic-waste-population-total-subset}
ggplot(data = plastic_waste_subset,
       aes(x = plastic_waste_per_cap,
       y = total_pop)) +
  geom_point()
```

```{r plastic-waste-population-coastal}
ggplot(data = plastic_waste,
       aes(x = plastic_waste_per_cap,
       y = coastal_pop)) +
  geom_point()
```

```{r plastic-waste-population-coastal-subset}
ggplot(data = plastic_waste_subset,
       aes(x = plastic_waste_per_cap,
       y = coastal_pop)) +
  geom_point()
```

It seems like there is a stronger linear association between plastic waste per capita and coastal population. 


### Exercise 5

Remove this text, and add your answer for Exercise 8 here.

```{r recreate-viz}
# insert code here
```


